---
title: "DublinPriceRegister"
output: html_document
date: "2023-02-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Dublin House Prices from 2010
Since 2010 Ireland's house property prices have been registered at https://www.propertypriceregister.ie. This module downloads the prices, filters to show only Dublin prices and stores as a tibble. 


```{r LoadLibraries, message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
```
Download the zip file from https://www.propertypriceregister.ie and unzip.

```{r propertypriceregister, message=FALSE, warning=FALSE, include=FALSE}
# Get the Property prices for Ireland
PriceUrl <- "https://www.propertypriceregister.ie/website/npsra/ppr/npsra-ppr.nsf/Downloads/PPR-ALL.zip/$FILE/PPR-ALL.zip"
download.file(PriceUrl, "PPR-ALL.zip")
unzip("PPR-ALL.zip")
```
Use readr to read in the resulting CSV file PPR-All.csv and save as PPReg. 
```{r}
PPReg <- readr::read_csv("PPR-All.csv", col_names= c("DateOfSale","Address","County","Eircode","Price","NotFullMarketPrice","VATExclusive","Description","Size"), col_types = c("c","c","c","c","c","c","c","c","c"),
                  skip=1,locale=default_locale() )

head(PPReg)
```

Use dplyr to correct the sale date to a date and the price as a number. The **sub('<80>','',)** function replaces the euro symbol with a blank space and **gsub(',','')** is used to replace the , in the the number format to allow the character to be converted to a number.

```{r}
options(digits = 14)
PPReg <- PPReg %>% 
          dplyr::mutate(DateOfSale = as.Date(DateOfSale, "%d/%m/%Y")) %>%
          dplyr::mutate(Price =  as.numeric(gsub(',', '', sub('<80>', '', Price))) )

head(PPReg)
```

Filter the results to show only the Dublin prices.
```{r propertypriceregister2}
# Get the Dublin property prices
DubPPReg <- PPReg[PPReg$County == "Dublin",]

head(DubPPReg)
```
View the most expensive houses by County.
```{r}
PPReg %>%
  dplyr::group_by(County) %>%
  dplyr::summarize(Max = max(Price), Min = min(Price)) %>%
  dplyr::arrange(desc(Max))
```
```{r}
DubPPReg %>%
  dplyr::mutate(Year = format(as.Date(DateOfSale, format="%Y-/%m-%d"),"%Y")) %>%
  dplyr::group_by(Year) %>%
  dplyr::summarize(Max = max(Price), Min = min(Price)) %>%
  dplyr::arrange(Year)
```
```{r}
rm(PriceUrl) # cleaning up unneccessary variables
```

