---
title: "Dublin House Prices"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{css, echo=FALSE}
table {
  margin: 0;
  border: .5px solid #666;
}
th {
  font-size: 14px; padding: 5px; border-bottom: 1px solid #ddd;font-weight: bold;
}
td {
  font-size: 14px; padding: 5px;
}

```
## Dublin House Prices from 2010
Since 2010 Ireland's house property prices have been registered at [propertypriceregister.ie](https://www.propertypriceregister.ie). This sample code downloads the house prices and stores as a tibble. 


```{r LoadLibraries, message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
library(scales)
library(zoo)
library(ggplot2)
```



Download the zip file from https://www.propertypriceregister.ie and unzip.

```{r propertypriceregister, message=FALSE, warning=FALSE, include=FALSE}
# Get the Property prices for Ireland
PriceUrl <- "https://www.propertypriceregister.ie/website/npsra/ppr/npsra-ppr.nsf/Downloads/PPR-ALL.zip/$FILE/PPR-ALL.zip"
download.file(PriceUrl, "PPR-ALL.zip")
unzip("PPR-ALL.zip")
```
Use the `read_csv` function in `readr` to read in the resulting CSV file PPR-All.csv and save as a tibble PPReg. 
```{r}
PPReg <- readr::read_csv("PPR-All.csv", col_names= c("Date_Of_Sale","Address","County","Eircode","Price","Not_Full_Market_Price","VAT_Exclusive","Description","Size"), col_types = c("c","c","c","c","c","c","c","c","c"),
                  skip=1,locale=default_locale() )

head(PPReg)
```

The Date_Of_Sale is converted from a character to a date using the `as.Date` function. 
To convert the price from a character to a number, first the number is *cleaned* by running the `sub('<80>','',)` function to remove the euro symbol and `gsub(',','')` to remove the commas, then the `as.numeric` function is run. 
```{r}
PPReg <- PPReg %>% 
         dplyr::mutate(Date_Of_Sale = as.Date(Date_Of_Sale, "%d/%m/%Y") ) %>%
          dplyr::mutate(Price = (as.numeric(gsub(',', '', sub('<80>', '', Price)))) ) 

head(PPReg)
```

Filter the results to get only the Dublin prices. Save the results as *DubPPReg*. 
```{r propertypriceregister2}
DubPPReg <- PPReg[PPReg$County == "Dublin",]
```


```{r}
options(knitr.kable.NA = "")
knitr::kable(DubPPReg[1:3, c(1,2,3,5,8)], format = "html", table.attr='class="PaddedTable"', digits = 2, format.args = list(decimal.mark = ".", big.mark = ",", scientific = FALSE), col.names = c("Date Of Sale",	"Address",	"County",	"Price",	"Description") )
```



## Summary of prices by County
View the highest, lowest and median house prices by county.
```{r}
CountySummary <- PPReg %>%
                  dplyr::group_by(County) %>%
                  dplyr::summarize(Median = round(median(Price)), Maximum = round(max(Price)), Lowest = round(min(Price)) ) %>%
                  dplyr::arrange(desc(Median)) 

knitr::kable(CountySummary, format = "html", digits = 2, format.args = list(decimal.mark = ".", big.mark = ",", scientific = FALSE))

```

### Summary of prices for Dublin by Year
View the highest, lowest and median house prices in Dublin by year

```{r}
DubByYear <- DubPPReg %>%
              dplyr::mutate(Year = format(as.Date(Date_Of_Sale, format="%Y-/%m-%d"),"%Y")) %>%
              dplyr::group_by(Year) %>%
              dplyr::summarize(Median = round(median(Price)), Maximum = round(max(Price)), Lowest = round(min(Price)) )  %>%
              dplyr::arrange(desc(Year)) 


knitr::kable(DubByYear, format = "html", digits = 2, format.args = list(decimal.mark = ".", big.mark = ",", scientific = FALSE))
```
```{r}
DubByYearMonth <- DubPPReg %>%
              dplyr::mutate(YearMonth = format(as.Date(Date_Of_Sale, format="%Y-/%m-%d"),"%Y%m")) %>%
              dplyr::mutate(YearQuarter = zoo::as.yearqtr(Date_Of_Sale, format = "%Y-%m-%d") ) %>%
              dplyr::group_by(YearQuarter) %>%
              dplyr::summarize(Median = round(median(Price)), Maximum = round(max(Price)), Lowest = round(min(Price)) )  %>%
              dplyr::arrange(desc(YearQuarter)) 
```


```{r}
  ggplot(DubByYearMonth, aes(YearQuarter, Median)) +
  labs(x = "", y = "") +
  scale_y_continuous(labels=scales::dollar_format(prefix = "€"))+
  theme(axis.text.x=element_text(angle=-45, hjust=0.001)) +
  scale_x_continuous(breaks=min(DubByYearMonth$YearQuarter):max(DubByYearMonth$YearQuarter)) +
  geom_line()
```

<p>
<p>  
### Remove unused variables
The zip file and csv file are removed to save space. Unneeded variables are removed using the `rm` function. 
```{r}
file_names <- c("PPR-All.csv","PPR-All.zip" )
unlink(file_names)
rm(PriceUrl) # cleaning up variables

```